MetaTracts - A Method for Robust Extraction and Visualization of Carbon Fiber Bundles in Fiber Reinforced Composites

### 2014-09-25 chapters/MetaTracts.tex
small fixes

Answering the above queries from the volume renderings of XCT datasets, or from visual inspection of particular 2D slices is non-trivial, even for experts. We detail and evaluate a method which uses visualization techniques to gain insight into our data. We demarcate the above queries into two parts: \textit{geometric structure} and \textit{spatial context}. Geometric structure refers to the shape, size, and orientation of a fiber bundle or a group of bundles. Spatial context refers to how two or more bundles interact with each other. 
We advance and interpret techniques from diffusion tensor imaging to extract and visualize geometric structures from 3D X-ray computed tomography data of the woven carbon fiber reinforced composites. The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in \textit{lower resolutions} and thus \textit{larger areas} of woven fabrics.
\begin{figure}[tb]
	\centering
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
	%\includegraphics[width=0.5\textwidth,  trim = 0mm 300mm 0mm 0mm, clip,]{imagesMT2014/image_algo}
	\includegraphics[width=0.475\textwidth]{images_pvis/workflow.pdf}
	\caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
	\label{fig:flowchart}
\end{figure}
Fig.~\ref{fig:flowchart} shows an over view of our approach. Starting from the XCT data, we first generate MetaTracts, cluster the MetaTracts to create fiber bundles and then generate visualizations from the clusters. For a detailed description of the MetaTracts generation process, see Sec.~\ref{subsec:MetaTracts-generation}. For the description of plausible fiber bundle generation see Sec.~\ref{subsec:fiber-bundles}. For a discussion of the visualizations we create for our domain experts see Sec.~\ref{sec:vis}. Sec.~\ref{sec:results} describes our experimental results, Sec.~\ref{sec:param_choices} presents our parameter choices and  Sec.~\ref{sec:user_eval} summarizes our user evaluations.

@@

\begin{figure}[tb]
\centering
%\includegraphics[width=0.5\textwidth, trim = 0mm 110mm 0mm 0mm, clip,]{images_pvis/figure1}
\includegraphics[width=0.45\textwidth]{images_pvis/data-char.pdf}
\caption{Data characteristics: (a) Rendering of the data. (b) 2D slices along the X-axis and (c) 2D slice along the Z-axis. (d) Magnified region marked in green square (c). Multiple fiber bundles cross and are indistinguishable by visual inspection alone. }
\label{fig:data-char}
\end{figure}
\section {Data Characteristic and Assumptions}
\label{sec:char_data}
Fig. \ref{fig:data-char} shows dataset 1 with woven fiber bundles. Dataset 1 is 450$\times$300$\times$500 voxels in size with isotropic resolution of 2$\mu m$. The data type is unsigned 8 bit integer. Fig.~\ref{fig:data-char} clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern. The unit cell is used as a basis structure for CFRP manufacturing. The alignment and weaving pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material. Fig.~\ref{fig:data-char}a shows a volume rendering of the dataset. Fig.~\ref{fig:data-char}b shows 2D slices along the $X$- and Fig.~\ref{fig:data-char}c in $Z$-axis. Fig.~\ref{fig:data-char}d shows the magnified version of the green region of interest. 
The resolution of dataset 1 renders the individual carbon fibers in a fiber bundle hard to resolve in the 2D slice images. Fig.~\ref{fig:data-char}d contains two bundles going in opposite directions. The separation between two fiber bundles is barely visible. The carbon fibers themselves do not differ much from the underlying epoxy matrix in terms of attenuation which poses an additional challenge for characterization. The fiber bundles may differ in terms of the amount of fibers in the bundle. Fig.~\ref{fig:data-char}a shows the large variation in cross section sizes among the bundles.
Depending on the weaving pattern, the fiber bundles cross each other in different orientations. In general, the number of orientations is limited by the weaving pattern. Nevertheless, the weaving pattern may cause the individual bundles to be curved. In consequence individual fibers may be adjacent in Euclidean distance but belong to different bundles. 
We make the following assumptions on our data:
\begin{itemize}[noitemsep]
\item The structure embedded in the data contains fiber bundles of indiscernible fibers.
\item Local orientation: each point within a fiber bundle has a local orientation which is parallel to the corresponding area in the fiber bundle.
\item The local orientation may change gradually inside the fiber bundle.
\item The local orientation may be noisy and not reliable.
\item Connectivity: moving along the direction of a non-noisy local orientation in small increments, we will reach another neighborhood with similar local orientation.
\item Fiber bundles going in different directions only interact near the surface of contact.
\end{itemize}
%\begin{figure}[htp]
%  \centering
%    \subfloat[Overview of the dataset]{\label{figur:3}\includegraphics[scale=0.3]{imagesMetaTracts/crop13-a.eps}}
%  \subfloat[X Axis]{\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/xaxis.eps}}\label{figur:1}\\
%  \subfloat[Y Axis]{\label{figur:2}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/yaxis.eps}}
%  \subfloat[Z Axis]{\label{figur:3}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/zaxis.eps}}
%  \caption{One of our datasets.Fig \ref{fig:orig_data}(a) provides an overview of the entire volume, the rest show slices of each of the major axis.}\label{fig:orig_data}
%\end{figure}

@@

%
\begin{figure}[tb]
\centering
%\includegraphics[width=0.45\textwidth]{images_pvis/figure3}
%\includegraphics[width=0.5\textwidth,  trim = 0mm 90mm 0mm 0mm, clip]{images_pvis/figure3}
\includegraphics[width=0.45\textwidth]{images_pvis/reliable_hessian.pdf}
\caption{Reliable Hessians. (a) Colored according to Orientation vector mapped to RGB. (b) 2D slice along Z and (c) along X axis. (d) Magnified region marked in green square (in c).}
\label{fig:reliable_hessian}
\end{figure}

@@

\subsection {MetaTracts generation}
\label{subsec:MetaTracts-generation}
 In this section we explain the process of MetaTracts generation in $\mathbb{R}^2$.The procedure trivially extends to $\mathbb{R}^3$. In $\mathbb{R}^2$, all the above properties hold except that cylinders are replaced by rectangles. The reliable Hessian step  generates a local orientation in $\mathbb{R}^2$ and a reliability measure $R_H$[0,1]. In Fig. \ref{fig:algo}, all regions which have unreliable Hessian have been marked in blue. Fig.~\ref{fig:algo}c shows a single cylinder of a MetaTracts. Let the seed point associated with the MetaTract be grid point $C_{p}$ (property 2 in Sec.~\ref{subsec:mtprop}). The local orientation at $C_p$ as computed in the step above is $N_P$ and is given by the dark green arrow. We generate a rectangle of length $L$ and radius $R$, the green dashed rectangle shows the rectangle generated at $C_p$ (properties 3 and 5 in Section~\ref{subsec:mtprop}). 
%
\begin{figure}[tb]
  \begin{center}
 \includegraphics[ width=\linewidth]{images_pvis/mt_gen.pdf}
  \end{center}
  \vspace{-1em}
 \caption{MetaTracts generation: (B) MetaTract generation, region with $R_H$ less than threshold. (A) close up of marked region in (B). (C) Parts of a single MetaTract.}
  \label{fig:algo}
    \vspace{-1.5em}
\end{figure}
The set of vertices which are in the green region but not overlapped by blue are possible candidates for the start point of the next cylinder in a particular MetaTract. We call these vertices \textit{``candidate vertices"}. From these we select another grid vertex which will be the start point for the next cylinder.
%\begin{figure}
%\centering
% \includegraphics[width=0.25\textwidth]{imagesMetaTracts/angle-diff.png}
%\caption{Angle between the local orientation of two cylinders that are part of the same MetaTract.}
%\label{fig:angle-diff}
%\end{figure}
Based on such assumptions on our data as local orientation and connectivity (Sec.~\ref{sec:char_data}), the order of the candidate vertices is based on the following characteristics:

@@

\begin{figure}[tb]
\centering
	\includegraphics[width=0.45\textwidth]{images_pvis/figure5.png}
	%\includegraphics[width=0.45\textwidth,  trim = 0mm 100mm 0mm 50mm, clip]{imagesMT2014/image_length}
	\caption{(A) MetaTracts color-coded according to their mean orientations. (B) length distribution of individual MetaTracts for a particular bundle (unit for length is grid cube edge length:  $2\mu m$). A single fiber bundle is made of MetaTracts of varying length}
	\label{fig:length_distribution}
\end{figure}

@@

\begin{figure}[tb] 
  \centering  	
  \includegraphics[width=\linewidth]{images_pvis/clustering.pdf}
  %\includegraphics[width=0.45\textwidth]{imagesMT2014/image_clustering}
  	\caption{(B) Result of K means clustering with data points projected to the top three eigenvectors as major axes. (A) MetaTracts belonging to orientation cluster 1. (C) MetaTracts belonging to orientation cluster 2. MetaTracts in gray(A,C) show context.
  	(D,E) shows the result of distance based clustering on the orientation cluster. (F) shows the combined result. }
  \label{fig:orientation_clustering}
  \end{figure}

@@

\subsection{Choice of clustering techniques}
As opposed to the two phase clustering we experimented with using proximity alone and using single step clustering of MetaTracts.
%\begin{figure}[tb]
%\centering
%		\includegraphics[width=0.4\textwidth]{images_pvis/figure7}
%		%\includegraphics[width=0.5\textwidth, trim = 0mm 70mm 0mm 20mm, clip]{imagesMT2014/image7b.pdf}
%	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
%	%\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
%	\caption{(A) Volume rendering of data set Fig.~\ref{fig:data-char}A.(B) shows a single slice of the data set. (C) shows two of the extracted meshes together. (D) shows the meshes separately.}
%	\label{fig:crop-16-decomp}
%\end{figure}  
%  
In comparison Fig.~\ref{fig:comparison} shows the MetaTracts in Fig.~\ref{fig:length_distribution}a being hierarchically clustered directly using proximity alone into 10, 15, and 20 clusters (ground truth is 11 clusters) without first performing orientation clustering. As discussed above, single linkage hierarchical clustering in the presence of overlapping fibers tends to create large clusters and small, low-cardinality outliers. When number of clusters ($h$) is 10 two large incorrect clusters are generated and the rest are outliers. As $h$ increases, some appropriate bundles start to form. Even at $h=20$, the top bundles all clusters together into one large incorrectly identified bundle.
%
On the other hand, Fig..~\ref{fig:comparison_orient} shows the MetaTracts in Fig..~\ref{fig:length_distribution}a, being clustered using K-means clustering after being embedded  in a m-dimensional (lower) space using proximity alone, without a second hierarchical clustering step. We set ``k" to be  10, 15, 20 clusters. In the lower dimensional space the spatial context is lost and fibers which are in reality far away are grouped together.
\begin{figure}[hb]
	\centering
	\includegraphics[width=0.45\linewidth]{images_pvis/comparison_all.pdf}
	\caption{Applying only hierarchical (a) and K-means (b) clustering methods of MetaTracts for various numbers of clusters.}
	\label{fig:comparison}
\end{figure}

@@

\begin{figure}
\centering
		\includegraphics[width=0.45\textwidth]{images_pvis/figure7}
		%\includegraphics[width=0.5\textwidth, trim = 0mm 70mm 0mm 20mm, clip]{imagesMT2014/image7b.pdf}
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
	%\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
	\caption{(A) Volume rendering of data set Fig.~\ref{fig:data-char}A.(B) shows a single slice of the data set. (C) shows two of the extracted meshes together. (D) shows the meshes separately.}
	\label{fig:crop-16-decomp}
\end{figure}

@@

\begin{figure}[tb]
\centering
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/image8b}
	\includegraphics[width=0.45\linewidth]{images_pvis/dataset2.pdf}
	\caption{Data set with flat thin and compact bundles. (A) shows the volume rendering and a 2D slice with one of the boundaries marked in green, (B) shows the clusters according to individual orientation. (C) shows the complete result. (D) shows the voxelization of (C).}
	\label{fig:prepreg}
\end{figure}

@@

\section {User Evaluation}
\label {sec:user_eval}
\begin{figure}[tb]
	\centering
	\includegraphics[width=0.45\textwidth,  trim = 0mm 00mm 0mm 0mm, clip]{images_pvis/usereval2}
	%\includegraphics[width=0.5\textwidth,  trim = 10mm 40mm 0mm 15mm, clip]{imagesMT2014/image_length_histogram.pdf}
	\caption{The user evaluation consists of a set of queries on videos of our results, volume rendering and Hessian colormap. The questionnaire tests the effectiveness of MetaTracts in visualizing ``geometric structure" and ``spatial context". The Likert scale goes from 1-5: strongly disagree, disagree, neutral, agree and strongly agree. }
	\label{fig:userstudy}
\end{figure}
Four NDT practitioners who are familiar with X-Ray computed tomography and its visual representations, four material scientists who are experienced in analyzing fiber reinforced polymers and four practitioners of visualization participated in the user evaluation of our work. We therefore setup a questionnaire with questions on solving given tasks as well as questions using Likert scale (LS) for evaluating the performance of the MetaTracts over currently used techniques. LS consists of 5 levels from ``Strongly disagree" to ``Strongly agree". Fig.~\ref{fig:comparison} shows the results. Task 1 of the questionnaire was to ``identify the number of fiber bundles" from a video showing a 360 degree rotation of the volume rendering of dataset 1 (Fig.~\ref{fig:data-char}A). This task tests the perception of geometric structure and how common problems like uniform gray scale values and occlusion of the different fiber bundles affects the user. As expected, Fig.~\ref{fig:comparison} shows that the participants found it difficult to determine the number of bundles from volume rendering alone. Another common approach is to color map the top eigenvector of Hessian matrix to RGB space. We show a video showing a 360 degree rotation of the volume rendering of the color mapped eigenvectors (similar to Fig.~\ref{fig:reliable_hessian}A) and asked the same query. Compared to the volume rendering, the Hessian color map provides more information about the major orientations and we see that the evaluators performed better in this test than in the direct volume rendering test before. The majority of our participants agreed or strongly agreed (LS) that Hessian color map was more effective. Finally, we showed the MetaTracts extracted bundles (for orientation 1 and 2 separately Fig.~\ref{fig:orientation_clustering}D,E). All participants could correctly identify the number of fiber bundles. All participants agreed that MetaTracts was more effective in perceiving geometric structure than direct volume rendering. They especially liked the 2D slices of the MetaTracts (Fig.~\ref{fig:crop-16-decomp}B, compared to Fig.~\ref{fig:data-char}C).
The next set of tasks on the MetaTracts results tests the spatial context perceived from MetaTracts. We achieve this by asking the evaluators to select the largest fiber bundle from a subset of fiber bundles, to select the smallest from a subset and given a fiber bundle and to select the most similar one from a subset. Due to the problems of occlusions and similar gray scale values these tests are difficult on volume rendering of the plain data only, but straight forward using MetaTracts. In the Likert scale, the majority strongly agreed that the tasks are undemanding compared to volume rendering.
Finally, the majority participants agreed that the extracted meshes provides better spatial context than volume rendering.

>>>>

Answering the above queries from the volume renderings of XCT datasets, or from visual inspection of particular 2D slices is non-trivial, even for experts. We detail and evaluate a method which uses visualization techniques to gain insight into our data. We demarcate the above queries into two parts: \textit{geometric structure} and \textit{spatial context}. Geometric structure refers to the shape, size, and orientation of a fiber bundle or a group of bundles. Spatial context refers to how two or more bundles interact with each other. 
We advance and interpret techniques from diffusion tensor imaging to extract and visualize geometric structures from 3D X-ray computed tomography data of the woven carbon fiber reinforced composites. The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in \textit{lower resolutions} and thus \textit{larger areas} of woven fabrics.
\begin{figure}[tb]
	\centering
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
	%\includegraphics[width=0.5\textwidth,  trim = 0mm 300mm 0mm 0mm, clip,]{imagesMT2014/image_algo}
	\includegraphics[width=\linewidth]{images_pvis/workflow.pdf}
	\caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
	\label{fig:flowchart}
\end{figure}
Fig.~\ref{fig:flowchart} shows an over view of our approach. Starting from the XCT data, we first generate MetaTracts, cluster the MetaTracts to create fiber bundles and then generate visualizations from the clusters. For a detailed description of the MetaTracts generation process, see Sec.~\ref{subsec:MetaTracts-generation}. For the description of plausible fiber bundle generation see Sec.~\ref{subsec:fiber-bundles}. For a discussion of the visualizations we create for our domain experts see Sec.~\ref{sec:vis}. Sec.~\ref{sec:results} describes our experimental results, Sec.~\ref{sec:param_choices} presents our parameter choices and  Sec.~\ref{sec:user_eval} summarizes our user evaluations.

@@

\begin{figure}[tb]
\centering
%\includegraphics[width=0.5\textwidth, trim = 0mm 110mm 0mm 0mm, clip,]{images_pvis/figure1}
\includegraphics[width=\linewidth]{images_pvis/data-char.pdf}
\caption{Data characteristics: (a) Rendering of the data. (b) 2D slices along the X-axis and (c) 2D slice along the Z-axis. (d) Magnified region marked in green square (c). Multiple fiber bundles cross and are indistinguishable by visual inspection alone. }
\label{fig:data-char}
\end{figure}
\section {Data Characteristic and Assumptions}
\label{sec:char_data}
Fig. \ref{fig:data-char} shows dataset 1 with woven fiber bundles. Dataset 1 is 450$\times$300$\times$500 voxels in size with isotropic resolution of 2$\mu m$. The data type is unsigned 8 bit integer. Fig.~\ref{fig:data-char} clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern. The unit cell is used as a basis structure for CFRP manufacturing. The alignment and weaving pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material. Fig.~\ref{fig:data-char}a shows a volume rendering of the dataset. Fig.~\ref{fig:data-char}b shows 2D slices along the $X$- and Fig.~\ref{fig:data-char}c in $Z$-axis. Fig.~\ref{fig:data-char}d shows the magnified version of the green region of interest. 
The resolution of dataset 1 renders the individual carbon fibers in a fiber bundle hard to resolve in the 2D slice images. Fig.~\ref{fig:data-char}d contains two bundles going in opposite directions. The separation between two fiber bundles is barely visible. The carbon fibers themselves do not differ much from the underlying epoxy matrix in terms of attenuation which poses an additional challenge for characterization. The fiber bundles may differ in terms of the amount of fibers in the bundle. Fig.~\ref{fig:data-char}a shows the large variation in cross section sizes among the bundles.
Depending on the weaving pattern, the fiber bundles cross each other in different orientations. In general, the number of orientations is limited by the weaving pattern. Nevertheless, the weaving pattern may cause the individual bundles to be curved. In consequence individual fibers may be adjacent in Euclidean distance but belong to different bundles. 
We make the following assumptions on our data:
\begin{itemize}[noitemsep]
\item The structure embedded in the data contains fiber bundles of indiscernible fibers.
\item Local orientation: each point within a fiber bundle has a local orientation which is parallel to the corresponding area in the fiber bundle.
\item The local orientation may change gradually inside the fiber bundle.
\item The local orientation may be noisy and not reliable.
\item Connectivity: moving along the direction of a non-noisy local orientation in small increments, we will reach another neighborhood with similar local orientation.
\item Fiber bundles going in different directions only interact near the surface of contact.
\end{itemize}
%\begin{figure}[htp]
%  \centering
%    \subfloat[Overview of the dataset]{\label{figur:3}\includegraphics[scale=0.3]{imagesMetaTracts/crop13-a.eps}}
%  \subfloat[X Axis]{\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/xaxis.eps}}\label{figur:1}\\
%  \subfloat[Y Axis]{\label{figur:2}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/yaxis.eps}}
%  \subfloat[Z Axis]{\label{figur:3}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/zaxis.eps}}
%  \caption{One of our datasets.Fig \ref{fig:orig_data}(a) provides an overview of the entire volume, the rest show slices of each of the major axis.}\label{fig:orig_data}
%\end{figure}

@@

%
\begin{figure}[tb]
\centering
%\includegraphics[width=0.45\textwidth]{images_pvis/figure3}
%\includegraphics[width=0.5\textwidth,  trim = 0mm 90mm 0mm 0mm, clip]{images_pvis/figure3}
\includegraphics[width=\linewidth]{images_pvis/reliable_hessian.pdf}
\caption{Reliable Hessians. (a) Colored according to Orientation vector mapped to RGB. (b) 2D slice along Z and (c) along X axis. (d) Magnified region marked in green square.}
\label{fig:reliable_hessian}
\end{figure}

@@

\subsection {MetaTracts generation}
\label{subsec:MetaTracts-generation}
 In this section we explain the process of MetaTracts generation in $\mathbb{R}^2$.The procedure trivially extends to $\mathbb{R}^3$. In $\mathbb{R}^2$, all the above properties hold except that cylinders are replaced by rectangles. The reliable Hessian step  generates a local orientation in $\mathbb{R}^2$ and a reliability measure $R_H$[0,1]. In Fig. \ref{fig:algo}, all regions which have unreliable Hessian have been marked in blue. Fig.~\ref{fig:algo}c shows a single cylinder of a MetaTracts. Let the seed point associated with the MetaTract be grid point $C_{p}$ (property 2 in Sec.~\ref{subsec:mtprop}). The local orientation at $C_p$ as computed in the step above is $N_P$ and is given by the dark green arrow. We generate a rectangle of length $L$ and radius $R$, the green dashed rectangle shows the rectangle generated at $C_p$ (properties 3 and 5 in Section~\ref{subsec:mtprop}). 
%
\begin{figure}[tb]
  \begin{center}
 \includegraphics[ width=\linewidth]{images_pvis/mt_gen.pdf}
  \end{center}
  \vspace{-1em}
 \caption{MetaTracts generation: (b) MetaTract generation, region with $R_H$ less than threshold. (a) close up of marked region in (b). (c) Parts of a single MetaTract.}
  \label{fig:algo}
    \vspace{-1.5em}
\end{figure}
The set of vertices which are in the green region but not overlapped by blue are possible candidates for the start point of the next cylinder in a particular MetaTract. We call these vertices \textit{``candidate vertices"}. From these we select another grid vertex which will be the start point for the next cylinder.
%\begin{figure}
%\centering
% \includegraphics[width=0.25\textwidth]{imagesMetaTracts/angle-diff.png}
%\caption{Angle between the local orientation of two cylinders that are part of the same MetaTract.}
%\label{fig:angle-diff}
%\end{figure}
Based on such assumptions on our data as local orientation and connectivity (Sec.~\ref{sec:char_data}), the order of the candidate vertices is based on the following characteristics:

@@

\begin{figure}[tb]
\centering
	\includegraphics[width=\linewidth]{images_pvis/figure5.png}
	%\includegraphics[width=0.45\textwidth,  trim = 0mm 100mm 0mm 50mm, clip]{imagesMT2014/image_length}
	\caption{(a) MetaTracts color-coded according to their mean orientations. (b) length distribution of individual MetaTracts for a particular bundle (unit for length is grid cube edge length:  $2\mu m$). A single fiber bundle is made of MetaTracts of varying length}
	\label{fig:length_distribution}
\end{figure}

@@

\begin{figure}[tb] 
  \centering  	
  \includegraphics[width=\linewidth]{images_pvis/clustering.pdf}
  %\includegraphics[width=0.45\textwidth]{imagesMT2014/image_clustering}
  	\caption{(b) Result of K means clustering with data points projected to the top three eigenvectors as major axes. (a) MetaTracts belonging to orientation cluster 1. (c) MetaTracts belonging to orientation cluster 2. MetaTracts in gray (a,c) show context.
  	(d,e) shows the result of distance based clustering on the orientation cluster. (f) shows the combined result. }
  \label{fig:orientation_clustering}
  \end{figure}

@@

\subsection{Choice of clustering techniques}
As opposed to the two phase clustering we experimented with using proximity alone and using single step clustering of MetaTracts.
%\begin{figure}[tb]
%\centering
%		\includegraphics[width=0.4\textwidth]{images_pvis/figure7}
%		%\includegraphics[width=0.5\textwidth, trim = 0mm 70mm 0mm 20mm, clip]{imagesMT2014/image7b.pdf}
%	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
%	%\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
%	\caption{(A) Volume rendering of data set Fig.~\ref{fig:data-char}A.(B) shows a single slice of the data set. (C) shows two of the extracted meshes together. (D) shows the meshes separately.}
%	\label{fig:crop-16-decomp}
%\end{figure}  
%  
Figure~\ref{fig:comparison}a shows the MetaTracts that are hierarchically clustered directly using proximity alone into 10, 15, and 20 clusters (ground truth is 11 clusters) without first performing orientation clustering. As discussed above, single linkage hierarchical clustering in the presence of overlapping fibers tends to create large clusters and small, low-cardinality outliers. When number of clusters ($h$) is 10 two large incorrect clusters are generated and the rest are outliers. As $h$ increases, some appropriate bundles start to form. Even at $h=20$, the top bundles all clusters together into one large incorrectly identified bundle.
%
On the other hand, Figure.~\ref{fig:comparison}b shows the MetaTracts, clustered using K-means clustering after being embedded in a $m$-dimensional (lower) space using proximity alone, without a second hierarchical clustering step for 10, 15, and 20 clusters. In the lower dimensional space the spatial context is lost and fibers which are in reality far away are grouped together.
\begin{figure}[tb]
	\centering
	\includegraphics[width=\linewidth]{images_pvis/comparison_all.pdf}
	\caption{Applying only hierarchical (a) and K-means (b) clustering methods of MetaTracts for various numbers of clusters.}
	\label{fig:comparison}
\end{figure}

@@

\begin{figure}
\centering
		\includegraphics[width=\linewidth]{images_pvis/figure7}
		%\includegraphics[width=0.5\textwidth, trim = 0mm 70mm 0mm 20mm, clip]{imagesMT2014/image7b.pdf}
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
	%\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
	\caption{(a) Voxelization of data set 1; (b) a single slice of the data set; (c) two of the extracted meshes together; (d) the meshes rendered separately.}
	\label{fig:crop-16-decomp}
\end{figure}

@@

\begin{figure}[tb]
\centering
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/image8b}
	\includegraphics[width=\linewidth]{images_pvis/dataset2.pdf}
	\caption{Data set with flat thin and compact bundles. (a) shows the volume rendering and a 2D slice with one of the boundaries marked in green, (b) shows the clusters according to individual orientation. (c) shows the complete result. (d) shows the voxelization of (c).}
	\label{fig:prepreg}
\end{figure}

@@

\section {User Evaluation}
\label {sec:user_eval}
\begin{figure}[tb]
	\centering
	\includegraphics[width=\linewidth,  trim = 0mm 00mm 0mm 0mm, clip]{images_pvis/usereval2}
	%\includegraphics[width=0.5\textwidth,  trim = 10mm 40mm 0mm 15mm, clip]{imagesMT2014/image_length_histogram.pdf}
	\caption{The user evaluation consists of a set of queries on videos of our results, volume rendering and Hessian colormap. The questionnaire tests the effectiveness of MetaTracts in visualizing ``geometric structure" and ``spatial context". The Likert scale goes from 1-5: strongly disagree, disagree, neutral, agree and strongly agree. }
	\label{fig:userstudy}
\end{figure}
Four NDT practitioners who are familiar with X-Ray computed tomography and its visual representations, four material scientists who are experienced in analyzing fiber reinforced polymers and four practitioners of visualization participated in the user evaluation of our work. We therefore setup a questionnaire with questions on solving given tasks as well as questions using Likert scale (LS) for evaluating the performance of the MetaTracts over currently used techniques. LS consists of 5 levels from ``Strongly disagree" to ``Strongly agree". Fig.~\ref{fig:comparison} shows the results. Task 1 of the questionnaire was to ``identify the number of fiber bundles" from a video showing a 360 degree rotation of the volume rendering of dataset 1 (Fig.~\ref{fig:data-char}A). This task tests the perception of geometric structure and how common problems like uniform gray scale values and occlusion of the different fiber bundles affects the user. As expected, Fig.~\ref{fig:comparison} shows that the participants found it difficult to determine the number of bundles from volume rendering alone. Another common approach is to color map the top eigenvector of Hessian matrix to RGB space. We show a video showing a 360 degree rotation of the volume rendering of the color mapped eigenvectors (similar to Fig.~\ref{fig:reliable_hessian}A) and asked the same query. Compared to the volume rendering, the Hessian color map provides more information about the major orientations and we see that the evaluators performed better in this test than in the direct volume rendering test before. The majority of our participants agreed or strongly agreed (LS) that Hessian color map was more effective. Finally, we showed the MetaTracts extracted bundles (for orientation 1 and 2 separately Fig.~\ref{fig:orientation_clustering}D,E). All participants could correctly identify the number of fiber bundles. All participants agreed that MetaTracts was more effective in perceiving geometric structure than direct volume rendering. They especially liked the 2D slices of the MetaTracts (Fig.~\ref{fig:crop-16-decomp}B, compared to Fig.~\ref{fig:data-char}C).
The next set of tasks on the MetaTracts results tests the spatial context perceived from MetaTracts. We achieve this by asking the evaluators to select the largest fiber bundle from a subset of fiber bundles, to select the smallest from a subset and given a fiber bundle and to select the most similar one from a subset. Due to the problems of occlusions and similar gray scale values these tests are difficult on volume rendering of the plain data only, but straight forward using MetaTracts. In the Likert scale, the majority strongly agreed that the tasks are undemanding compared to volume rendering.
Finally, the majority participants agreed that the extracted meshes provides better spatial context than volume rendering.

\section{Parameter choices}
\label{sec:param_choices}
\begin{figure}[tb]
\centering
	\includegraphics[width=\linewidth,  trim = 0mm 260mm 0mm 15mm, clip]{images_pvis/figure9}
	%\includegraphics[width=0.5\textwidth,  trim = 10mm 40mm 0mm 15mm, clip]{imagesMT2014/image_length_histogram.pdf}
	\caption{Number of tracts and median, minimum and maximum length of individual MetaTracts in the orientation cluster Fig.~\ref{fig:orientation_clustering}A, clustered into 10 clusters. The unit for length is the grid cube length. Fig.~\ref{fig:length_distribution}(B) shows the length of individual MetaTracts in cluster 6. Clusters labeled 2,3,4,5 and 7 all have cardinality less than 10. }
	\label{fig:len_dist_crop16}
\end{figure}
The critical parameters are $K$ for the K-means in orientation based clustering and $h$ for the hierarchical clustering. K denotes the number of major directions, which is either known a priori or can be easily estimated by looking at the weaving pattern. 
Our framework is robust to the choice of $h$ (Sec.~\ref{sec:results}). Large $h$ keeps the major bundles intact. For example, in  Fig.~\ref{fig:len_dist_crop16}, where $h$ was set to ten while the ground truth was five clusters, we observe that the major clusters remain well segmented and small clusters can be removed if they have too few elements (clusters 2,3,4,5 and 7 in Fig.~\ref{fig:len_dist_crop16} all have less than 10 elements). This is an appealing trait of the proximity based clustering, thus providing good results even when exact $h$ might be unknown.
The following parameters were fixed for all the tests. We set the reliable Hessian threshold $R_{h}$ to be 0.3 for all our tests. A $R_{h}$ of 0.0 would mean all points have reliable local orientation which would cause spurious MetaTracts detection. On the other hand, a very high $ R_{h}$ would lead to be a decline in number of MetaTracts produced. The $\alpha$ and $\beta$ in $R_{h}$ are as explained in Frangi et al. \cite{Frangi1998} and set to 0.5. The length and the width parameters for the cylinders for the MetaTracts decide how coarse the fiber cylinders are. Larger cylinders will handle noisy local orientation better as it inspects a higher number of candidate points to extend the fiber. We used 10.00 and 2.00 for length and breadth(measured in grid cube length), respectively for all tests. Short tracts of less than 0.01 times the smallest axes size are prone to have erroneous local orientation and are therefore removed. $\alpha$ and $\beta$ in equation~\ref{eqn:algo_1} decide how quickly the value of the factor decays; we have used [7-10] and half the length of cylinder (5.0). Our number of fiber bundle directions are limited. Thus even for small $m$ (cardinality of lower dimension in orientation clustering), the distinction between the orientation clusters is preserved quite well. We compared  $m=3-7$ experimentally without any dramatic change in results. 
%\begin{figure}[tb]
%	\centering
%	\includegraphics[width=0.5\textwidth, trim = 0mm 80mm 0mm 170mm, clip]{images_pvis/figure10}
%	\caption{Direct hierarchical clustering of MetaTracts Fig.~\ref{fig:length_distribution}(A) by proximity into 10,15 and 20 clusters.}
%	\label{fig:comparison}
%\end{figure}
%
